### 2019-03-22 IEEE_TC_submission.d/symmetry_properties.tex
ssr

@@

The canalyzing variable in each of the first $n-2$ lines (one per
layer) can be any variable that has not yet occurred in a preceding
line.  The last layer contains the remaining two variables.  Thus,
there are $n! /[n-(n-2)]!$ = $n!/2$ possibilities for the pattern of variable
occurrences in $f$.

>>>>

\subsection{Testing Whether an $r$-Symmetric Function is an NCF}
\label{sse:rsym_to_ncf}

Theorem~\ref{thm:ncf_r_symmetric} shows that given an NCF $f$,
the problem of finding the smallest integer $r$ such that
$f$ is $r$-symmetric can be solved efficiently.
We now show that the converse problem, that is, determining whether a given
$r$-symmetric Boolean function is an NCF,
can also be solved efficiently.
When the answer is ``yes", our algorithm also constructs a
default-normalized representation of $f$.

Let $f(x_1, x_2, \ldots, x_n)$ be an $r$-symmetric Boolean function of $n$
Boolean variables.
We assume that the symmetry groups $g_i$, $1 \leq i \leq r$,
of $f$ are given.
Let $m_i$ denote the number of variables
in the group containing variable $x_i$, $1 \leq i \leq m$.
Thus, in the group containing $x_i$, the number
of variables which can take on the value 1 varies from 0
to $m_i$, $1 \leq i \leq n$.
We also assume that $f$ is
given as a table $T$ with $\mu = \prod_{i=1}^r (m_i+1)$ rows.
Each row of $T$ specifies an $r$-tuple $(c_1, c_2, \ldots, c_r)$,
where $c_i$ is the number of variables of group $g_i$ which
have the value 1, along with the \{0,1\} value of $f$ for that $r$-tuple.
As will be seen, our algorithm for determining whether $f$ is
an NCF runs in $O(r^2\mu)$ time.
The algorithm relies on the following observation.

\begin{observation}\label{obs:rsym_canalyzing}
Suppose $f(x_1, x_2, \ldots, x_n)$ is an $r$-symmetric Boolean function.
Consider any variable $x_i$ and suppose
the number of variables in the group
containing $x_i$ is $m_i$, $1 \leq i \leq r$.
\begin{enumerate}
\item The function $f$ is canalyzing in variable $x_i$
with canalyzing value 1 iff all the table entries where the number
of 1's in the group containing $x_i$ is nonzero have the same value
of $f$. (This value of $f$ is the canalyzed value when $x_i = 1$.)
\item The function $f$ is canalyzing in variable $x_i$ with
canalyzing value 0 iff all the table entries where the number of
1's in the group containing $x_i$ is less than $m_i$ have the same
value of $f$. (This value of $f$ is the canalyzed value when $x_i = 0$.)
\end{enumerate}
\end{observation}
We now specify an iterative algorithm for determining whether 
$f$ is an NCF, and if so, constructing a simplified 
NCF representation (defined in Section~\ref{sec:intro}) for $f$.
In iteration $j$, the algorithm uses a set $X_j$ of remaining variables,
table $T_j$ with $\mu_j$ rows, and the number of groups $g_j$.
Initially (i.e., $j = 1$), $X_1$ = $\{x_1, \ldots, x_n\}$,
$T_1$ is the given table $T$ for $f$, 
$\mu_1 = \mu$ (the number of rows of $T$) and $g_j = r$ (the number of
symmetry groups of $f$) and the NCF representation for $f$ is empty.
The algorithm executes as long as one or more symmetry groups remain.
In each iteration, the algorithm carries out the following steps.

\smallskip
\noindent
I. Use Observation~\ref{obs:rsym_canalyzing} to determine whether there is
a variable $x_i \in X_j$ is canalyzing.

\smallskip
\noindent
II. If yes, perform the following steps. 
\begin{enumerate}
\item Let $\alpha$ and $\beta$ be the respective canalyzing and 
canalyzed values for $x_i$ found in Step~I. 
\item For each variable $x_p$ in the group containing $x_i$, add the rule~
``$x_p~:~\alpha ~\longrightarrow~ \beta$"~ to the NCF representation of $f$.
\item Obtain $X_{j+1}$ by deleting from $X_j$ all the variables which are 
in the same group as $x_i$.
\item Set $g_{j+1} = g_j - 1$.
\item Obtain $T_{j+1}$ by retaining only those rows of $T_j$ where all the
variables in the group containing $x_i$ are set to $\overline{\alpha}$. 
\item Start the next iteration.
\end{enumerate}

\smallskip
\noindent
III. (Here, Step~I didn't find a canalyzing variable.)~ 
Output ``$f$ is not an NCF" and \textbf{stop}.

\smallskip
\noindent
The correctness of the algorithm is a direct consequence of
Observation~\ref{obs:rsym_canalyzing}. If the algorithm is successful,
it produces a simplified NCF representation of $f$; this can then be
efficiently converted into the default-normalized representation as
discussed in Section~\ref{sse:ncf_layer}.

To estimate the running time, we note that in iteration $j$, we can
determine in $O(g_j\,\mu_j)$ = $O(r\,\mu)$ time using 
Observation~\ref{obs:rsym_canalyzing}
whether there is a variable that is canalyzing.
(This can be done by considering one arbitrary variable 
from each group.) 
Once a variable is identified, the other steps in that iteration
can be completed in $O(\mu_j)$ = $O(\mu)$ time.
Thus, the time for each iteration is $O(r\,\mu)$.
Since there are at most $r$ iterations, the running time of the
algorithm is $O(r^2\,\mu)$.
The following theorem summarizes this result.

\begin{theorem}\label{thm:rsym_canalyzing}
Given an $r$-symmetric function $f$, the problem of testing whether
$f$ is an NCF can be solved efficiently. \QED
\end{theorem}

@@

The canalyzing variable in each of the first $n-2$ lines (one per
layer) can be any variable that has not yet occurred in a preceding
line.  The last layer contains the remaining two variables.  Thus,
there are $n! /[n-(n-2)]!$ = $n!/2$ possibilities for the pattern of 
canalyzing variable occurrences in $f$.
